Seeing Double: Transient Dips and Photometric Binarity in PTFO$\,$8-8695

### 2020-04-23 paper/ms.tex
improve text

\begin{abstract}
  PTFO$\,$8-8695b is a candidate hot Jupiter in the 7--10 million year
  old Orion-OB1a cluster. We inspected data from TESS and Gaia to
  clarify whether it is truly a planet.  The Gaia data show that
  PTFO$\,$8-8695 is a photometric binary with respect to members of
  its kinematic group, and also suggest that it is an astrometric
  binary.  The TESS lightcurve shows that the dominant variability in
  this system is a sinusoid with a long period $P_{\rm
  \ell}=11.96\,$hr, presumably caused by stellar rotation.  Also
  present is a combined dip and asymmetric sinusoid, previously
  identified as the planet candidate, that repeats with a short period
  $P_{\rm s}= 10.74\,$hr.  The two signals beat every 4.48 days.
  Although there is a dip in the short-period signal, ground-based
  photometry from the past decade indicates that the orbital phase of
  the dip seems to have instantaneously jumped, at least once, and
  perhaps twice.  Planets do not ``jump'' in orbital phase.  Given the
  available evidence, we believe that PTFO$\,$8-8695 is a binary M
  dwarf in which one star shows the long rotation signal, and the
  other star shows ``transient dipping'' that has been seen in a few
  other young low-mass stars.  The origin of these transient dips is
  still undetermined.
\end{abstract}

@@

There are perhaps a few challenges to the planetary interpretation (if
these ``features'' are not already seen as such).  They include that
the planet does not seem to emit infrared radiation in occulation, at
least at the expected amplitude \citep{yu_tests_2015}.  In
addition, despite measurement attempts by multiple investigators, it
does not seem to show the Rossiter effect at the amplitude expected
given the rapid stellar rotation and large planet size
\citep{yu_tests_2015,ciardi_followup_2015}.  Finally, detailed
modelling of the ``precession + gravity darkening'' transits has shown
that the necessary degree of gravity darkening is too great, given the
spectroscopically observed equatorial velocity
\citep{howarth_reappraisal_2016}.  Additionally, as the
gravity-darkened star precessed about its rotation axis, it would need
to show photometric variability that has not been observed.

While the planetary interpetation clearly faces challenges,
alternative explanations do as well.  High-latitude accretion hotspots
might produce the observed H$\alpha$ variability, and require
fine-tuning produce dips of the approach duration. Furthermore, \ptfo\
does not have an infrared (IR) excess associated with the presence of
an inner disk \citep[{\it e.g.},][Figure~18]{yu_tests_2015}.
Low-latitude starspots, hot or cold, struggle to produce the necessary
dip durations.

Another potential explanation is that between 0.1\% and 1\% of rapidly
rotating low-mass stars in $\mathcal{O}$(10)$\,$Myr old associations
show narrow dips in phase with strong stellar rotation signals
\citep{rebull_usco_2018}.  The dips can persist over months, but their
depths often vary, and sometimes change immediately after stellar
flares.  The explanation proposed by \citet{stauffer_orbiting_2017}
and \citet{david_transient_2017} to explain this novel class of
variable stars is that a circumstellar cloud of dust or gas might be
orbiting near the Keplerian co-rotation radius.  Given available
ground-based data, this explanation is not clearly viable for \ptfo,
because a clear determination of the stellar rotation period has not
yet been possible
\citep{van_eyken_ptf_2012,koen_multicolour_2015,raetz_yeti_2016}.

We begin in Section~\ref{sec:observations} by describing newly
available observations from TESS \citep{ricker_transiting_2015} and
Gaia \citep{gaia_collaboration_gaia_2018}.  The TESS lightcurve shows
two distinct signals, which we extract and analyze in
Section~\ref{sec:tess}.  A long-period sinusoid repeats every 12.0
hours, and is probably stellar rotation.  A short-period dip superposed on an
asymmetric sinusoid repeats every 10.7 hours.  Analyzing the
Gaia data in Section~\ref{sec:gaia}, we show that relative to its
kinematic group, \ptfo\ is a photometric binary, and suggest it also an
astrometric binary.  We collect and discuss the puzzle pieces in
Section~\ref{sec:discussion}.  The orbital phase of the dip seems to
have instantaneously changed over the past decade.  In addition, a
number of other young stars show lightcurve morphologies similar to
the short-period signal.  We therefore argue that PTFO$\,$8-8695 is a
binary M dwarf in which one star shows a rotation signal, and the
other shows ``transient dipping'' potentially caused by eclipses of plasma at
the Keplerian co-rotation radius.  Section~\ref{sec:conclusions}
summarizes our main points.

@@

The 2-minute stacks for \ptfo\ were reduced to lightcurves by the
Science Processing Operations Center (SPOC) at NASA
Ames~\citep{jenkins_tess_2016}.  Our main analysis used the resulting
Presearch Data Conditioning (PDC) lightcurve.  The PDC lightcurve
aperture used pixels chosen to maximize the SNR of the total flux of
the target \citep{smith_kepler_apertures_2017}.  Non-astrophysical
variability was removed through the methods discussed by
\citet{smith_kepler_PDC_2017}.

@@

As a precursor to more detailed analysis, we calculated generalized
Lomb-Scargle periodograms using \texttt{astrobase}
\citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015,bhatti_astrobase_2018}.
The two largest peaks in the Lomb-Scargle periodogram of the
lightcurve were clearly separated at a ``short'' period $P_{\rm s}
\approx 0.448\,{\rm days}$ and a ``long'' period $P_{\rm \ell} \approx
0.499\,{\rm days}$.  The $P_{\rm \ell}$ peak had the greater power of
the two.  Smaller harmonics surrounding each of these two dominants
peaks were also present.

@@

Initial signal-processing experiments fitting out splines or sinusoids
showed that after subtracting out the long-period signal, the
short-period signal dominated the periodogram, and vice-versa.
However it quickly became clear that it would be beneficial to
simultaneously model the signals separately, in order to preserve the
power at each frequency.

@@

We fitted the lightcurve as a linear combination of Fourier harmonics
at the short and long periods, plus a transit at the short period.
Symbolically, the total flux $f$ is given as
\begin{equation}
  f = f_{\rm s} + f_{\rm \ell}
  = f_{\rm transit,s} + f_{\rm Fourier,s} + f_{\rm Fourier,\ell},
\end{equation}
where $f_{\rm s}$ is the relative flux at the short period, and
$f_{\rm \ell}$ is the flux at the long period.  Writing out the
Fourier terms,
\begin{align}
  f = &f_{\rm transit,s} + \sum_{n=1}^{N} A_n \sin(n\omega_{\rm s}t)
  + \sum_{n=1}^{N} B_n \cos(n\omega_{\rm s}t)\\
  &+ \sum_{m=1}^{M} A_m \sin(m[\omega_{\rm \ell}t+\phi_{\rm \ell}])
  + \sum_{m=1}^{M} B_m \cos(m[\omega_{\rm \ell}t+\phi_{\rm \ell}]), \nonumber
\end{align}
for $N$ and $M$ the total number of harmonics at the short and long
periods, respectively, $A_i$ and $B_i$ the amplitudes for each
harmonic term (potentially negative), and $\omega_i = 2\pi / P_i$ the
angular frequency for $i$ the short or long period index.  We fixed
the ``phase-offset'' for the short period signal to be zero, and let
the reference time for the long period signal float by introducing
$\phi_{\rm \ell}$.  Since we did not a priori know how many harmonics
would be appropriate, we considered a number of different choices for
$N$ and $M$, and used the Bayesian information criterion to choose the
appropriate model (Table~1).

@@

		\vspace{-0.7cm}
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f5b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Evidence for binarity in \ptfo}.
    {\it Top: Hertzsprung-Russell diagram of \ptfo\ and late-type
    members of 25$\,$Ori-1.} Members of the 25$\,$Ori-1 group (black
    circles) were identified by \citet{kounkel_apogee2_2018} through
    clustering on six-dimensional Gaia DR2 and APOGEE-2 data.  The
    ``neighborhood'' (gray circles) is non-member stars within 5
    standard deviations of the mean 25$\,$Ori-1 right ascension,
    declination, and parallax.  It contains members of the Orion
    complex with its full spread of ages, in addition to field
    interlopers.  $G$ denotes Gaia broadband magnitudes, $Bp$ Gaia
    blue, $Rp$ Gaia red, and $\omega_{\rm as}$ the parallax in
    arcseconds.  The $x$-axis limits have been set to show only K and
    M dwarf members, to accentuate \ptfo's separation from the
    single-star sequence.  {\it Bottom: Astrometric goodness-of-fit
    versus $Rp$ magnitude for 25$\,$Ori-1 members}.  The single-source
    astrometric model for \ptfo\ provides a poor fit, which could be due to
    stellar variability or binarity.  Cluster members that
    are at least as variable as \ptfo\ show lower astrometric excesses
    (black squares), suggesting binarity as the root cause.
		\label{fig:gaia}
	}
\end{figure*}

@@

\subsection{Long period sinusoid}

@@

\subsection{Short period dip}

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.5\textwidth]{f6.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Timing residuals for \ptfob\ from a decade of monitoring.}
		Black points are times of dips, minus the indicated linear
		ephemeris.  The $y$-axis is given in units of phase for the
		short-period signal.  The star shows the binned TESS ephemeris.
		Dips have been observed by \citet{van_eyken_ptf_2012},
		\citet{ciardi_followup_2015}, \citet{yu_tests_2015},
		\citet{raetz_yeti_2016}, \citet{onitsuka_multicolor_2017}, and
		\citet{tanimoto_evidence_2020}.  Certain dips ({\it e.g.}, the one
		at phase 0 in mid-2019) are consistent with noise, and were likely
		reported because something was {\it expected}, rather than
		convincingly {\it observed}.  Horizontal dashed lines are drawn at
		$\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$, highlighting either a
		numerical coincidence or an observational bias.  The orbital phase
		observed by TESS (lower-right) is consistent with that of
		\citet{tanimoto_evidence_2020}, and quite different from the
		original phase.
		\label{fig:o_minus_c}
	}
\end{figure}

@@

\subsection{Short period out-of-dip modulation}

@@

\subsubsection{Planetary interpretation}
If there were a giant planet transiting \ptfo, it would tidally
distort the host star, and cause ellipsoidal photometric modulations
that also peak at quadrature \citep[see][]{shporer_astrophysics_2017}.
Interpreting the second cosine harmonic as planet-induced tidal distortion,
it would imply a minimum planet mass $M_{\rm p} \sin i$ of
$3.8\,M_{\rm Jup}$.  For this estimate, we assumed $R_\star = 1.39
R_\odot$, and $M_\star = 0.39 M_\odot$ \citep{van_eyken_ptf_2012}.
This ellipsoidal amplitude is larger than the typical
modulations induced by close-in giant planets because the host star is
puffy, and still on the pre-main-sequence.

@@

\begin{figure*}[hbtp]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f7.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf \ptfo\ and its brethren.}
    Five transient and persistent flux dip stars selected based on
    their visual similarity to the short-period signal in \ptfo are as
    shown.  They include EPIC 204143627, EPIC 204321142, EPIC
    204270520, EPIC 205483258 (RIK-210), and EPIC 204787516.  RIK-210
    has the longest period of any of these objects.  EPIC 204787516
    has a less angular dip, and could also be an eclipsing binary.
    All stars have similar properties as \ptfo, but are in Upper~Sco.
    We found these objects through studies by
    \citet{stauffer_orbiting_2017}, \citet{david_transient_2017}, and
    \citet{rebull_usco_2018}.
		\label{fig:brethren}
	}
\end{figure*}

@@

These lightcurves have been phenomenologically classified as
``persistent flux dips'' or ``transient flux dips'', based on whether
their depths and durations show variability over the 90-day K2
campaigns \citep{stauffer_orbiting_2017}.
As \citet{stauffer_orbiting_2017} explain,
these  objects are morphologically distinct from ``scallop shell' lightcurves,
and are present in stars at more advanced evolutionary disk stages than
the ``dipper'' stars \citep{ansdell_young_2016,cody_manyfaceted_2018}.
The persistent and transient flux dip stars have the following properties.
\begin{enumerate}[topsep=0.5ex,itemsep=-0.5ex,partopsep=1ex,parsep=1.5ex]
  \item Weak-lined T Tauri stars.
  \item Spectral type typically M2-M5 ({\it e.g.},
    \citealt{rebull_usco_2018},~Figure~20).
  \item Ages $\lesssim$ 100$\,$Myr.
  \item Stars typically lack spectroscopic accretion indicators 
  \item Phased lightcurves show shallow, angular dips, usually
    superposed on large-amplitude smooth variability. The latter
    is interpreted as stellar rotation.
  \item Rotation is fast---usually between 0.5 and 2.0 days.
  \item Rarely any infrared excess (never any W4 detection; only a few
    cases with W3 detctions).
  \item Can show multiple dips per cycle.
  %({\it e.g.}, EPIC 203692610, RIK-210).
  \item Dip depths, durations, and phases can vary over just a few cycles
    ({\it e.g.}, EPIC 204143627).
  \item Dip depths can change after flares.
  \item Rare at a population level: $\lesssim 1\%$ of relevant stars
    \citep{rebull_usco_2018}.
\end{enumerate}
The 10.74 hour signal in \ptfo\ meets all of these criteria.
This is the first connection of \ptfo\ with this class of objects
because the TESS lightcurve was needed to resolve the different
rotation signals.

@@

Second, between 40-50\% of the transient flux dip stars discovered in
$\rho$~Oph and Upper~Sco show two
Lomb-Scargle periods, and so are apparently binaries
\citep[][Table~1]{stauffer_orbiting_2017}.
This is higher than the corresponding main-sequence companion
fraction of ${\rm CF}_{0.1\text{-}0.5\,M_\odot}^{\rm MS} = 33 \pm 5\%$
\citep{henry_solar_2006,duchene_stellar_2013}.
Low-mass pre-main-sequence stars however have been shown to 
companion fractions up to twice as high in dispersed clusters such as
Upper~Sco and Taurus \citep{kraus_mapping_2008,kraus_mapping_2011}.
A detailed high-resolution imaging survey would be necessary to determine
whether the transient flux dip stars truly have any distinct
population-level binarity properties relative to other young low-mass
stars.

@@

The main physical question remaining is clearly what causes the
transient dipping. This is an unsolved problem not only for \ptfo\ but
also for an emerging class of similar young rapidly rotating M-dwarfs.
Many possible explanations been detailed by \citet{rebull_rotation_2016}, \citet{david_transient_2017},
\citet{stauffer_orbiting_2017}, and \citet{zhan_complex_2019}.
Their disfavored explanations include that the dips are caused by
{\it (i)} eclipsing binaries;
{\it (ii)} ``dipper''-flavor Class-I or Class-II disks;
{\it (iii)} eclipses of prominences;
{\it (iv)} high-latitude accretion hotspots;
{\it (v)} high-latitude starspots;
or
{\it (vi)} dust clouds of plausible composition.
We also view the possibility of {\it (vii)} tidally disrupted planetary
or cometary material to be implausible, given the synchronicity
between dip and rotation periods seen across many systems.

@@

The combination of TESS and Gaia observations have provided new
clarity in the \ptfo\ system.  Our main results are as follows.
\begin{itemize}
  \item {\it The TESS light-curve shows two periodic signals.} The
    ``long'' signal is a 10\% peak-to-peak sinusoid that repeats every
    11.96 hours.  The ``short'' signal is a 4\% peak-to-peak ``dip +
    asymmetric sinusoid'' that repeats every 10.74 hours. The signals
    beat, and therefore cannot be an artifact linked to data
    processing.
    Within the angular resolution of the Gaia source
    catalog, both signals need to originate from \ptfo.
  \item {\it The Gaia data imply binarity.} Relative to stars in its
    kinematic group, \ptfo\ is a photometric binary
    (Figure~\ref{fig:gaia}, top).  Relative to stars in its group that
    are at least as photometrically variable, \ptfo\ also shows
    signs of astrometric binarity (Figure~\ref{fig:gaia}, bottom).
  \item {\it The orbital phase of the dip has changed since the
    discovery by \citet{van_eyken_ptf_2012}.} As shown in
    Figure~\ref{fig:o_minus_c}, the phase seems to have jumped,
    perhaps twice. This agrees with the recent study by
    \citet{tanimoto_evidence_2020}.
  \item {\it All properties of \ptfo\ are consistent with the emerging
    class of transient and persistent flux dip stars.} Analogous
    lightcurves are shown in Figure~\ref{fig:brethren}.  Properties of
    this variability class are enumerated in
    Section~\ref{subsec:dipstars}.
\end{itemize}

The physical mechanism that explains the transient and persistent flux
dips is unresolved. Our preferred explanations include
transiting clumps of gas at the Keplerian corotation radius, and
occultations of starspots by an optically thick disk \citep[{\it
e.g.},][]{stauffer_orbiting_2017,david_transient_2017,zhan_complex_2019}.
The jumping orbital phase disfavors the enticing explanation of an
enshrouded, transiting protoplanet.  Though \ptfob seems unlikely
to be a planet, understanding it and its analogs is a worthy
problem that promises to teach us about the natal environments of the
majority of habitable-zone Earth-sized planets in the Milky Way
\citep{dressing_occurrence_2013}.

>>>>

\begin{abstract}
  PTFO$\,$8-8695b is a candidate hot Jupiter in the 7--10 million year
  old Orion-OB1a cluster. We inspected data from TESS and Gaia to
  clarify whether it is truly a planet.  The Gaia data show that
  PTFO$\,$8-8695 is a photometric binary with respect to members of
  its kinematic group, and independently suggest that it is an
  astrometric binary.  The TESS lightcurve shows that the dominant
  variability in this system is sinusoidal with a long period $P_{\rm
  \ell}=11.96\,$hr, presumably caused by stellar rotation.  Also
  present is a combined dip and asymmetric sinusoid, previously
  identified as the planet candidate, that repeats with a short period
  $P_{\rm s}= 10.74\,$hr.  The two signals beat every 4.48 days.
  Although there is a dip in the short-period signal, ground-based
  photometry from the past decade indicates that the orbital phase of
  the dip seems to have instantaneously jumped, at least once, and
  perhaps twice.  Planets do not ``jump'' in orbital phase.  Given the
  available evidence, we believe that PTFO$\,$8-8695 is a binary M
  dwarf in which one star shows the long rotation signal, and the
  other star shows transient dipping that has been seen in a few other
  young low-mass stars.  The origin of these transient dips is still
  undetermined.
\end{abstract}

@@

There are perhaps a few challenges to the planetary interpretation (if
these ``features'' are not already seen as such).  They include that
the planet does not seem to emit infrared radiation in occulation, at
least at the expected amplitude \citep{yu_tests_2015}.  In addition,
despite measurement attempts by multiple investigators, \ptfob\ does
not seem to show the Rossiter effect at the amplitude expected given
the rapid stellar rotation and large planet size
\citep{yu_tests_2015,ciardi_followup_2015}.  Finally, detailed
modelling of the ``precession + gravity darkening'' transits has shown
that the necessary degree of gravity darkening is too great, given the
spectroscopically observed equatorial velocity
\citep{howarth_reappraisal_2016}.  Additionally, as the
gravity-darkened star precessed about its rotation axis, it would need
to show photometric variability that has not been observed.

While the planetary interpetation clearly faces challenges,
alternative explanations do as well.  High-latitude accretion hotspots
might produce the observed H$\alpha$ variability, but require
fine-tuning to produce dips of the appropriate duration. Furthermore,
\ptfo\ does not have an infrared (IR) excess associated with the
presence of dust at the inner edge of a primordial disk \citep[{\it
e.g.},][Figure~18]{yu_tests_2015}.  Low-latitude starspots, hot or
cold, struggle to produce the necessary dip durations.

One potentially relevant fact is that between 0.1\% and 1\% of rapidly
rotating low-mass stars in $\mathcal{O}$(10)$\,$Myr old associations
show narrow dips in phase with strong stellar rotation signals
\citep{rebull_usco_2018}.  The dips can persist over months, but their
depths often vary, and sometimes change immediately after stellar
flares.  The explanation proposed by \citet{stauffer_orbiting_2017}
and \citet{david_transient_2017} to explain this novel class of
variable stars is that a circumstellar cloud of gas might be orbiting
near the Keplerian co-rotation radius.  To date, this explanation has
not clearly been viable for \ptfo, because a clear determination of
the stellar rotation period has not yet been possible
\citep{van_eyken_ptf_2012,koen_multicolour_2015,raetz_yeti_2016}.

We begin in Section~\ref{sec:observations} by describing newly
available observations from TESS \citep{ricker_transiting_2015} and
Gaia \citep{gaia_collaboration_gaia_2018}.  The TESS lightcurve shows
two distinct signals, which we extract and analyze in
Section~\ref{sec:tess}.  A long-period sinusoid repeats every 12.0
hours, and is probably stellar rotation.  A short-period dip
superposed on an asymmetric sinusoid repeats every 10.7 hours.
Analyzing the Gaia data in Section~\ref{sec:gaia}, we show that
relative to its kinematic group, \ptfo\ is a photometric binary, and
suggest it also an astrometric binary.  We discuss the puzzle pieces
in Section~\ref{sec:discussion}.  The orbital phase of the dip seems
to have instantaneously changed over the past decade.  In addition,
other young stars show lightcurve morphologies similar to the
short-period signal.  We therefore argue that PTFO$\,$8-8695 is a
binary M dwarf in which one star shows a rotation signal, and the
other shows transient dipping potentially caused by eclipses of
plasma at the Keplerian co-rotation radius.
Section~\ref{sec:conclusions} summarizes our main points.

@@

The 2-minute stacks for \ptfo\ were reduced to lightcurves by the
Science Processing Operations Center (SPOC) at NASA
Ames~\citep{jenkins_tess_2016}.  Our main analysis used the resulting
Presearch Data Conditioning (PDC) lightcurve.  The PDC lightcurve
aperture used pixels chosen to maximize the SNR of the total flux of
the target \citep{smith_kepler_apertures_2017}.  Non-astrophysical
variability was removed by fitting out trends common to many stars
\citep{smith_kepler_PDC_2017}.

@@

As a precursor to more detailed analysis, we calculated generalized
Lomb-Scargle periodograms using \texttt{astrobase}
\citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015,bhatti_astrobase_2018}.
The two largest peaks in the Lomb-Scargle periodogram of the
lightcurve were separated at a ``short'' period $P_{\rm s}
\approx 0.448\,{\rm days}$ and a ``long'' period $P_{\rm \ell} \approx
0.499\,{\rm days}$.  The $P_{\rm \ell}$ peak had the greater power of
the two.  Smaller harmonics surrounding each of these two dominants
peaks were also present.

@@

Initial signal-processing experiments fitting out splines or sinusoids
showed that after subtracting out the long-period signal, the
short-period signal dominated the periodogram, and vice-versa.
However it quickly became clear that it would be beneficial to
simultaneously model each signal, in order to preserve the
power at each frequency.

@@

We fitted the lightcurve as a linear combination of Fourier harmonics
at the short and long periods, plus a transit at the short period.
Symbolically, the total flux $f$ is given as
\begin{equation}
  f = f_{\rm s} + f_{\rm \ell}
  = f_{\rm transit,s} + f_{\rm Fourier,s} + f_{\rm Fourier,\ell},
\end{equation}
where $f_{\rm s}$ is the flux at the short period, and
$f_{\rm \ell}$ is the flux at the long period.  Writing out the
Fourier terms,
\begin{align}
  f = &f_{\rm transit,s} + \sum_{n=1}^{N} A_n \sin(n\omega_{\rm s}t)
  + \sum_{n=1}^{N} B_n \cos(n\omega_{\rm s}t)\\
  &+ \sum_{m=1}^{M} A_m \sin(m[\omega_{\rm \ell}t+\phi_{\rm \ell}])
  + \sum_{m=1}^{M} B_m \cos(m[\omega_{\rm \ell}t+\phi_{\rm \ell}]), \nonumber
\end{align}
for $N$ and $M$ the total number of harmonics at the short and long
periods, respectively, $A_i$ and $B_i$ the amplitudes for each
harmonic term (potentially negative), and $\omega_i = 2\pi / P_i$ the
angular frequency for $i$ the short or long period index.  We fixed
the ``phase-offset'' for the short period signal to be zero, and let
the reference time for the long period signal float by introducing
$\phi_{\rm \ell}$.  Since we did not a priori know how many harmonics
would be appropriate, we considered a number of different choices for
$N$ and $M$, and used the Bayesian information criterion to choose the
appropriate model (Table~1).

@@

		\vspace{-0.7cm}
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f5b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Evidence for binarity in \ptfo}.
    {\it Top: Hertzsprung-Russell diagram of \ptfo\ and late-type
    members of 25$\,$Ori-1.} Members of the 25$\,$Ori-1 group (black
    circles) were identified by \citet{kounkel_apogee2_2018}.  The
    ``neighborhood'' (gray circles) is non-member stars within 5
    standard deviations of the mean 25$\,$Ori-1 right ascension,
    declination, and parallax.  It contains members of the Orion
    complex with its full spread of ages, in addition to field
    interlopers.  $G$ denotes Gaia broadband magnitudes, $Bp$ Gaia
    blue, $Rp$ Gaia red, and $\omega_{\rm as}$ the parallax in
    arcseconds.  The $x$-axis limits are set to show only K and
    M dwarfs, to accentuate \ptfo's separation from the
    single-star sequence.  {\it Bottom: Astrometric goodness-of-fit
    versus $Rp$ magnitude for 25$\,$Ori-1 members}.  The single-source
    astrometric model for \ptfo\ provides a poor fit, which could be due to
    stellar variability or binarity.  Cluster members that
    are at least as variable as \ptfo\ show lower astrometric excesses
    (black squares), suggesting binarity as the probable root cause.
		\label{fig:gaia}
	}
\end{figure*}

@@

\subsection{Long Period Sinusoid}

@@

\subsection{Short Period Dip}

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.5\textwidth]{f6.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Timing residuals for \ptfob\ from a decade of monitoring.}
		Black points are times of dips, minus the indicated linear
		ephemeris.  The $y$-axis is given in units of phase for the
		short-period signal.  The star shows the binned TESS ephemeris.
		Dips have been observed by \citet{van_eyken_ptf_2012},
		\citet{ciardi_followup_2015}, \citet{yu_tests_2015},
		\citet{raetz_yeti_2016}, \citet{onitsuka_multicolor_2017}, and
		\citet{tanimoto_evidence_2020}.  Certain dips ({\it e.g.}, the one
		at phase 0 in mid-2019) are consistent with noise, and were likely
		reported because dips were expected, rather than
		convincingly observed.  Horizontal dashed lines are drawn at
		$\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$, highlighting either a
		numerical coincidence or an observational bias.  The orbital phase
		observed by TESS (lower-right) is consistent with that of
		\citet{tanimoto_evidence_2020}.
		\label{fig:o_minus_c}
	}
\end{figure}

@@

\subsection{Short Period Out-of-Dip Modulation}

@@

\subsubsection{Ellipsoidal Variability?}
If there were a giant planet transiting \ptfo, it would tidally
distort the host star, and cause ellipsoidal photometric modulations
that also peak at quadrature \citep[see][]{shporer_astrophysics_2017}.
Interpreting the second cosine harmonic as planet-induced tidal distortion,
it would imply a minimum planet mass $M_{\rm p} \sin i$ of
$3.8\,M_{\rm Jup}$.  For this estimate, we assumed $R_\star = 1.39
R_\odot$, and $M_\star = 0.39 M_\odot$ \citep{van_eyken_ptf_2012}.
This ellipsoidal amplitude is larger than the typical
modulations induced by close-in giant planets because the host star is
puffy, and still on the pre-main-sequence.

@@

\begin{figure*}[hbtp]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f7.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf \ptfo\ and its brethren.}
    Five transient and persistent flux dip stars selected based on
    their visual similarity to the short-period signal in \ptfo are as
    shown.
    They include
    EPIC 204143627,
    EPIC 204270520,
    EPIC 204321142,
    EPIC 246938594,
    and
    EPIC 205483258 (RIK-210).
    %and EPIC 204787516.
    RIK-210 has the longest period of any of these objects. 
    All the analogs displayed are either in Taurus or Upper~Sco, and
    meet the characteristics of Section~\ref{subsec:dipstars}.
    % EPIC 204787516 has a less angular dip, and could also be an
    % eclipsing binary.
    % All stars have similar properties as \ptfo, but are in Upper~Sco.
    We found these objects through studies by
    \citet{stauffer_orbiting_2017}, \citet{david_transient_2017}, and
    \citet{rebull_usco_2018}.
		\label{fig:brethren}
	}
\end{figure*}

@@

These lightcurves have been phenomenologically classified as
``persistent flux dips'' or ``transient flux dips'', based on whether
their depths and durations show variability over the 90-day K2
campaigns \citep{stauffer_orbiting_2017}.
As \citet{stauffer_orbiting_2017} explain,
these  objects are morphologically distinct from ``scallop shell' lightcurves,
and are present in stars at more advanced evolutionary disk stages than
the ``dipper'' stars \citep{ansdell_young_2016,cody_manyfaceted_2018}.
The persistent and transient flux dip stars all show angular dips
that are cannot be explained with starspots.
They typically have the following properties.
\begin{enumerate}[topsep=0.5ex,itemsep=-0.5ex,partopsep=1ex,parsep=1.5ex]
  \item Weak-lined T Tauri stars.
  \item Spectral type typically M2-M5 ({\it e.g.},
    \citealt{rebull_usco_2018},~Figure~20).
  \item Ages $\lesssim$ 100$\,$Myr.
  \item Stars typically lack spectroscopic accretion indicators 
  \item Phased lightcurves show shallow, angular dips, usually
    superposed on large-amplitude smooth variability. The latter
    is interpreted as stellar rotation.
  \item Rotation is fast---usually between 0.5 and 2.0 days.
  \item Rarely any infrared excess (never any W4 detection; only a few
    cases with W3 detctions).
  \item Can show multiple dips per cycle.
  %({\it e.g.}, EPIC 203692610, RIK-210).
  \item Dip depths, durations, and phases can vary over just a few cycles
    ({\it e.g.}, EPIC 204143627).
  \item Dip depths can change after flares.
  \item Rare at a population level: $\lesssim 1\%$ of relevant stars
    \citep{rebull_usco_2018}.
\end{enumerate}
The 10.74 hour signal in \ptfo\ meets all of these criteria.
This is the first connection of \ptfo\ with this class of objects
because the TESS lightcurve was needed to resolve the different
rotation signals.

@@

Second, between 40-50\% of the transient flux dip stars discovered in
$\rho$~Oph and Upper~Sco show two
Lomb-Scargle periods, and so are apparently binaries
\citep[][Table~1]{stauffer_orbiting_2017}.
This is higher than the main-sequence companion
fraction of ${\rm CF}_{0.1\text{-}0.5\,M_\odot}^{\rm MS} = 33 \pm 5\%$
\citep{henry_solar_2006,duchene_stellar_2013}.
Low-mass pre-main-sequence stars however have been shown to 
companion fractions up to twice as high in dispersed clusters such as
Upper~Sco and Taurus \citep{kraus_mapping_2008,kraus_mapping_2011}.
A detailed high-resolution imaging survey would be necessary to determine
whether the transient flux dip stars truly have any distinct
population-level binarity properties relative to other young low-mass
stars.

@@

The main physical question is what causes the
transient dipping. This is an unsolved problem not only for \ptfo\ but
also for an emerging class of similar young rapidly rotating M-dwarfs.
Many possible explanations been detailed by \citet{rebull_rotation_2016}, \citet{david_transient_2017},
\citet{stauffer_orbiting_2017}, and \citet{zhan_complex_2019}.
Their disfavored explanations include that the dips are caused by
{\it (i)} eclipsing binaries;
{\it (ii)} ``dipper''-flavor Class-I or Class-II disks;
{\it (iii)} eclipses of prominences;
{\it (iv)} high-latitude accretion hotspots;
{\it (v)} high-latitude starspots;
or
{\it (vi)} dust clouds of plausible composition.
We also view the possibility of {\it (vii)} tidally disrupted planetary
or cometary material to be implausible, given the synchronicity
between dip and rotation periods seen across many systems.

@@

The combination of TESS and Gaia has provided new
clarity in the \ptfo\ system.  Our main results are as follows.
\begin{itemize}
  \item {\it The TESS light-curve shows two periodic signals.} The
    ``long'' signal is a 10\% peak-to-peak sinusoid that repeats every
    11.96 hours.  The ``short'' signal is a 4\% peak-to-peak ``dip +
    asymmetric sinusoid'' that repeats every 10.74 hours. The signals
    beat, and therefore cannot be an artifact linked to data
    processing.
    Within the angular resolution of the Gaia source
    catalog, both signals need to originate from \ptfo.
  \item {\it The Gaia data imply binarity.} Relative to stars in its
    kinematic group, \ptfo\ is a photometric binary
    (Figure~\ref{fig:gaia}, top).  Relative to stars in its group that
    are at least as photometrically variable, \ptfo\ also shows
    signs of astrometric binarity (Figure~\ref{fig:gaia}, bottom).
  \item {\it The orbital phase of the dip has changed since the
    discovery by \citet{van_eyken_ptf_2012}.} As shown in
    Figure~\ref{fig:o_minus_c}, the phase seems to have jumped,
    perhaps twice. This agrees with the recent study by
    \citet{tanimoto_evidence_2020}.
  \item {\it All properties of \ptfo\ are consistent with the emerging
    class of transient and persistent flux dip stars.} Analogous
    lightcurves are shown in Figure~\ref{fig:brethren}.  Properties of
    this variability class are enumerated in
    Section~\ref{subsec:dipstars}.
\end{itemize}

The physical mechanism that explains the transient and persistent flux
dips is unresolved. Our preferred explanations include
transiting clumps of gas at the Keplerian corotation radius, and
occultations of starspots by an optically thick disk \citep[{\it
e.g.},][]{stauffer_orbiting_2017,david_transient_2017,zhan_complex_2019}.
The jumping orbital phase disfavors the explanation of an
enshrouded, transiting protoplanet.  Though \ptfob\ seems unlikely
to be a planet, understanding it and its analogs is a worthy
problem that promises to teach us about the natal environments of the
majority of habitable-zone Earth-sized planets in the Milky Way
\citep{dressing_occurrence_2013}.
